### 2019-04-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Galère ce chapitre 5, je sais pas trop quoi faire.

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\section{Lower Bound: Proof of Theorem~\ref{thm:5:BetterLowerBound}}
\label{proof:5:BetterLowerBound}

@@

% -----------------------------------------------------------------
\subsection{Proof of Theorem~\ref{thm:5:BetterLowerBound}}

@@

\subsection{Proof of Lemma~\ref{lem:5:CD}}

@@

\section{Proofs Elements Related to Regret Upper Bounds}
\label{proof:5:RegretUpperBounds}

This Appendix includes the main proofs, missing from the content of the article,
that yield the regret upper bound.
We start by controlling the sub-optimal draws when the \klUCB{} indices are used (instead of \UCB),
with any of our proposed algorithms (\MCTopM, \RandTopM) or \rhoRand{}. Then we focus on controlling collisions for \MCTopM-\klUCB.

\subsection{Control of the Sub-optimal Draws for \klUCB: Proof of Lemma~\ref{lem:5:SubOptimalSelections}}
\label{proof:5:SubOptimalSelections}

%
%     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
%     As only user $j$ is considered here, the superscript $j$ is dropped.
%
Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
\begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.

Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
\begin{align*}
\E_{\mu}[T_k^j(T)]
&= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
= \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
%&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
%
\intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
\E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
&\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
&\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
%
&\leq
\sum_{t=1}^T
  \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
+
\sum_{m=1}^M
  \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
\end{align*}
where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.

Now each of the two terms in the right hand side can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
The rightmost term can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact 1 in Appendix A of \cite{KLUCBJournal}.
The leftmost term can be controlled using Lemma~\ref{lem:5:Fact2KLUCB} stated below, that is a direct consequence of the proof of Fact 2 in Appendix A of \cite{KLUCBJournal}.

\begin{lemma}\label{lem:5:Fact1KLUCB}
    For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
    \begin{equation}
      \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
    \end{equation}
\end{lemma}

Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).

\begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
\begin{align*}
  &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
  + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
\end{lemma}

Putting things together, one obtains the non-asymptotic upper bound
\begin{align}\label{eq:5:UBprecise}
  \E_{\mu}\left[T_k^j(T)\right]
  & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
  & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
\end{align}
which yields Lemma~\ref{lem:5:SubOptimalSelections},
with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.

% --------------------------------------------------------------------------------
\subsection{Proof of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}}\label{proof:5:elementaryLemma_RandTopM_MCTopM}

Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
\begin{align*}
  & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
  & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
  & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
\end{align*}

Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
\[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}. To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
\begin{align}
  & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
  =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
  & \hspace{1cm} \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
  & \hspace{1cm}  \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
\end{align}
where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
\[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]

If $\mu_k > \mu_{k'}$, we rather use that
\[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
\[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
Summing over $k'$ yields the result.

@@

% --------------------------------------------------------------------------------
\subsection{Controlling Collisions for \MCTopM: Proof of Lemma~\ref{lem:5:collisionsMCTopM}}
\label{proof:5:collisionsMCTopM}

A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small. Its proof is postponed to the end of this section.

%
Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
%
The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''. This gives a logarithmic number of transitions $(2)$ and $(3)$, and so gives $\E_{\mu}[\sum_k\cC_k(T)] = \bigO{\log T}$,
with explicit constants depending on $\boldsymbol{\mu}$ and $M$.

\begin{figure}[h!]
  \resizebox{0.98\textwidth}{!}{
  \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
      %
      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
      %
      \draw [black,->] (start) -> (notfixed.20);
      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
      %
  \end{tikzpicture}
  }
  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
  \label{fig:5:StateMachineAlgorithm_MCTopM}
\end{figure}

% Explanation of the figure

As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
% following the algorithm of \MCTopM:
\begin{equation}
    s^j(t+1) =
    \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
    \cap \left( A^j(t) \in \TopM(t) \right).
\end{equation}

For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words. At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
or she stays in state $\overline{s^j(t)}$,
with two cases.
Either she sampled $A^j(t+1)$ uniformly
from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
with transition $(3)$,
in case of collision and if $A^j(t+1) \in \TopM(t)$,
or she sampled $A^j(t+1)$ uniformly
from $\TopM(t)$ with transition $(2)$,
if $A^j(t+1) \notin \TopM(t)$.
In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
%
% The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
% and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
% and experienced a collision at time $t-1$
Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).

>>>>

% -----------------------------------------------------------------
% -----------------------------------------------------------------
\subsection{Lower Bound: Proof of Theorem~\ref{thm:5:BetterLowerBound}}
\label{proof:5:BetterLowerBound}

@@

% -----------------------------------------------------------------
\subsubsection{Proof of Theorem~\ref{thm:5:BetterLowerBound}}

@@

\subsubsection{Proof of Lemma~\ref{lem:5:CD}}

@@

% \section{Proofs Elements Related to Regret Upper Bounds}
% \label{proof:5:RegretUpperBounds}

% This Appendix includes the main proofs, missing from the content of the article,
% that yield the regret upper bound.
% We start by controlling the sub-optimal draws when the \klUCB{} indices are used (instead of \UCB),
% with any of our proposed algorithms (\MCTopM, \RandTopM) or \rhoRand{}. Then we focus on controlling collisions for \MCTopM-\klUCB.

% \subsection{Control of the Sub-optimal Draws for \klUCB: Proof of Lemma~\ref{lem:5:SubOptimalSelections}}
% \label{proof:5:SubOptimalSelections}

% %
% %     Recall $g_k^j(t) \in \mathbb{R}$ denote the index of arm $k$ for user $j$ at time $t$.
% %     As only user $j$ is considered here, the superscript $j$ is dropped.
% %
% Fix $k\in\Mworst$ and a player $j \in \{1,\dots,M\}$.
% The key observation is that for \MCTopM, \RandTopM{} as well as the \rhoRand{} algorithm, it holds that
% \begin{equation}\left(A^j(t) = k\right) = \left(A^j(t) = k , \exists m \in \Mbest : g_m^j(t) < g_k^j(t) \right).\label{eq:5:KeyInclusion}\end{equation}
% Indeed, for the three algorithms, an arm selected at time $t+1$ belongs to the set $\TopM(t)$ of arms with $M$ largest indices.
% If the sub-optimal arm $k$ is selected at time $t$, it implies that $k \in \TopM(t)$, and, because there are $M$ arms in both \Mbest{} and $\TopM(t)$, one of the arms in \Mbest{} must be excluded from $\TopM(t)$.
% In particular, the index of arm $k$ must be larger than the index of this particular arm $m$.

% Using \eqref{eq:5:KeyInclusion}, one can then upper bound the number of selections of arm $k$ by user $j$ up to round $T$ as
% \begin{align*}
% \E_{\mu}[T_k^j(T)]
% &= \E_{\mu}\left[ \sum_{t=1}^T \mathbbm{1}\left( A^j(t) = k \right) \right]
% = \sum_{t=1}^T \Pr\left( A^j(t) = k \right).\\
% %&= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\Mbest,\; g_m(t) < g_k(t) \right) \\
% &= \sum_{t=1}^T \Pr\left( A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}^j(t) < g_k^j(t) \right).
% %
% \intertext{Considering the relative position of the upper-confidence bound $g_{m^*}^j(t)$ and the corresponding mean $\mu_m^* = \mu_{m^*}$, one can write the decomposition}
% \E_{\mu}[T_k^j(T)] &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; g_{m^*}(t) \leq g_k(t) , \forall m \in \{1,\dots,M\}: \;  g_{m^*}(t) \geq \mu_m^* \right)\\
% &\;\;\;\;\;\;\;\; + \sum_{t=1}^T\Pr\left( \exists m \in\{1,\dots,M\}: \; g_{m^*}(t) < \mu_m^* \right) \\
% &\leq \sum_{t=1}^T\Pr\left(A^j(t) = k,\;\; \exists m \in\{1,\dots,M\}:\; \mu_{m}^* \leq g_k(t)\right)+ \sum_{m=1}^M\sum_{t=1}^T\Pr\left(g_{m^*}(t) < \mu_m^* \right) \\
% %
% &\leq
% \sum_{t=1}^T
%   \Pr\left( A^j(t) = k,\; \mu_{M^*} \leq g_k(t) \right)
% +
% \sum_{m=1}^M
%   \sum_{t=1}^T \Pr\left( g_{m^*}(t) < \mu_m^* \right),
% \end{align*}
% where the last inequality (for the first term) comes from the fact that $\mu_{M^*}$ is the smallest of the $\mu_{m^*}$ for $m \in \{1, \dots,M\}$.

% Now each of the two terms in the right hand side can directly be upper bounded using tools developed by \cite{KLUCBJournal} for the analysis of kl-UCB.
% The rightmost term can be controlled using Lemma~\ref{lem:5:Fact1KLUCB} below that relies on a self-normalized deviation inequality, whose proof exactly follows from the proof of Fact 1 in Appendix A of \cite{KLUCBJournal}.
% The leftmost term can be controlled using Lemma~\ref{lem:5:Fact2KLUCB} stated below, that is a direct consequence of the proof of Fact 2 in Appendix A of \cite{KLUCBJournal}.

% \begin{lemma}\label{lem:5:Fact1KLUCB}
%     For any arm $k$, if $g_k^{j}(t)$ is the \klUCB{} index with exploration function $f(t)=\log(t)+3\log\log(t)$,
%     \begin{equation}
%       \sum_{t=1}^T \Pr\left(g_k^{j}(t) < \mu_k\right) \leq 3 + 4e \log\log(T).
%     \end{equation}
% \end{lemma}

% Denote $\kl'(x,y)$ the derivative of the function $x \mapsto \kl(x,y)$ (for any fixed $y\neq 0, 1$).

% \begin{lemma}\label{lem:5:Fact2KLUCB} For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $g_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
% \begin{align*}
%   &\sum_{t=1}^T\Pr\left( A^j(t) = k, \mu_{k'} \leq g_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
%   + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.\end{align*}
% \end{lemma}

% Putting things together, one obtains the non-asymptotic upper bound
% \begin{align}\label{eq:5:UBprecise}
%   \E_{\mu}\left[T_k^j(T)\right]
%   & \leq \frac{\log(T) + 3 \log\log(T)}{\kl(\mu_k,\mu_{M^*})} + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{M^*})^2}{\kl(\mu_k,\mu_{M*})^3}}\sqrt{\log(T) + 3\log\log(T)}  \nonumber\\
%   & \;\;\;\;\;\;+ 2\left(\frac{\kl'(\mu_k,\mu_{M^*})}{\kl(\mu_k,\mu_{M^*})}\right)^2 + 4Me \log\log(T) + 3M+1,
% \end{align}
% which yields Lemma~\ref{lem:5:SubOptimalSelections},
% with explicit constants $C_{\boldsymbol{\mu}}$ and $D_{\boldsymbol{\mu}}$.

% % --------------------------------------------------------------------------------
% \subsection{Proof of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}}\label{proof:5:elementaryLemma_RandTopM_MCTopM}

% Using the behavior of the algorithm when the current arm leaves the set $\TopM$ (Line 4), one has
% \begin{align*}
%   & \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t)\right)  \\ & \leq \sum_{t=1}^T \bP\left(A^j(t) = k, k \notin \hat{M}^j(t), A^j(t+1) \in \TopM(t) \cap \{k': g_{k'}^j(t-1) \leq g_k^j(t-1)\}\right) \\
%   & \leq   \sum_{t=1}^T \sum_{k' \neq k}\bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right) \\
%   & = \sum_{k' \neq k} \underbrace{\sum_{t=1}^T \bP\left(A^j(t) = k,A^j(t+1) = k', g_{k'}^j(t) \geq g_{k}^j(t), g_{k'}^j(t-1) \leq g_{k}^j(t-1)  \right)}_{:= T_{k'}}
% \end{align*}

% Now, to control $T_{k'}$, we distinguish two cases. If $\mu_k < \mu_{k'}$, one can write
% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k'}^j(t) \leq \mu_{k'}\right) + \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)\]
% The first term in the right hand side is $o(\log(T))$ by Lemma~\ref{lem:5:Fact1KLUCB}. To control the second term, we apply the same trick that led to the proof of Lemma~\ref{lem:5:Fact2KLUCB} in \cite{KLUCBJournal}.
% Letting $\kl^+(x, y) := \kl(x, y) \indic(x \geq y)$,
% and $\widehat{\mu}^j_{k,s}$ be the empirical mean of the $s$ first observations from arm $k$ by player $j$, one has
% \begin{align}
%   & \sum_{t=1}^T \bP\left(A^j(t) = k, g_{k}^j(t-1) \geq \mu_{k'}\right)
%   =  \bE \left[ \sum_{t=1}^T\sum_{s=1}^{t-1} \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)}\indic{\left( s  \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(t)\right)} \right]\nonumber\\
%   & \hspace{1cm} \leq \bE \left[ \sum_{s=1}^{T}\indic{\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right)}\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} \right] \nonumber \\
%   & \hspace{1cm}  \leq \sum_{s=1}^T \bP\left( s \times \kl^+\left(\widehat{\mu}_{k,s}^j , \mu_k\right) \leq f(T)\right),\label{eq:5:FromHere}
% \end{align}
% where the last inequality uses that for all $s$, \[\sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t-1) = s\right)} = \sum_{t=s-1}^T \indic{\left(A^j(t) = k , N^j_k(t) = s + 1\right)} \leq 1.\]
% From \eqref{eq:5:FromHere}, the same upper bound as that of Lemma~\ref{lem:5:Fact2KLUCB} can be obtained using the tools from \cite{KLUCBJournal}, which proves that for $T\to\infty$,
% \[T_{k'} = \frac{\log(T)}{\kl(\mu_k,\mu_{k'})} + o(\log(T)).\]

% If $\mu_k > \mu_{k'}$, we rather use that
% \[T_{k'} \leq \sum_{t=1}^T \bP\left(g_{k}^j(t) \leq \mu_{k}\right) + \sum_{t=1}^T \bP\left(A^j(t+1) = k', g_{k'}^j(t) \geq \mu_{k}\right)\]
% and similarly Lemma~\ref{lem:5:Fact1KLUCB} and a slight variant of Lemma~\ref{lem:5:Fact2KLUCB} to deal with the modified time indices yields
% \[T_{k'} = \frac{\log(T)}{\kl(\mu_{k'},\mu_{k})} + o(\log(T)).\]
% Summing over $k'$ yields the result.

@@

% % --------------------------------------------------------------------------------
% \subsection{Controlling Collisions for \MCTopM: Proof of Lemma~\ref{lem:5:collisionsMCTopM}}
% \label{proof:5:collisionsMCTopM}

% A key feature of both the \RandTopM{} and \MCTopM{} algorithms is Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, that states that the probability of switching from some arm because this arm leaves $\TopM(t)$ is small. Its proof is postponed to the end of this section.

% %
% Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below provides a schematic representation of the execution of the \MCTopM{} algorithm, that has to be exploited in order to properly control the number of collisions.
% %
% The sketch of the proof is the following: by focusing only on collisions in the ``not fixed'' state, bounding the number of transitions $(2)$ and $(3)$ is enough.
% Then, we show that both the number of transitions $(3)$ and $(5)$ are small: as a consequence of Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}, the average number of these transitions is $\bigO{\log T}$.
% Finally, we use that the length of a sequence of consecutive transitions $(2)$ is also small (on average smaller than $M$), and except for possibly the first one, starting a new sequence implies a previous transition $(3)$ or $(5)$ to arrive in the state ``not fixed''. This gives a logarithmic number of transitions $(2)$ and $(3)$, and so gives $\E_{\mu}[\sum_k\cC_k(T)] = \bigO{\log T}$,
% with explicit constants depending on $\boldsymbol{\mu}$ and $M$.

% \begin{figure}[h!]
%   \resizebox{0.98\textwidth}{!}{
%   \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
%       %
%       \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
%       \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
%       \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
%       %
%       \draw [black,->] (start) -> (notfixed.20);
%       \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
%       \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
%       \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
%       \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
%       \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
%       %
%   \end{tikzpicture}
%   }
%   \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
%   Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
%   \label{fig:5:StateMachineAlgorithm_MCTopM}
% \end{figure}

% % Explanation of the figure

% As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
% Formally, $s^j(0)$ is false, and $s^j(t+1)$ is defined inductively from $s^j(t)$ as
% % following the algorithm of \MCTopM:
% \begin{equation}
%     s^j(t+1) =
%     \left( s^j(t) \cup \left( \overline{s^j(t)} \cap \overline{C^j(t)} \right) \right)
%     \cap \left( A^j(t) \in \TopM(t) \right).
% \end{equation}

% For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words. At step $t$, if player $j$ is not fixed ($\overline{s^j(t)}$), she can have three behaviors when executing \MCTopM.
% She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
% or she stays in state $\overline{s^j(t)}$,
% with two cases.
% Either she sampled $A^j(t+1)$ uniformly
% from $\TopM(t) \cap \{ m : g_m^j(t) \leq g_k^j(t) \}$
% with transition $(3)$,
% in case of collision and if $A^j(t+1) \in \TopM(t)$,
% or she sampled $A^j(t+1)$ uniformly
% from $\TopM(t)$ with transition $(2)$,
% if $A^j(t+1) \notin \TopM(t)$.
% In particular, note that if $\overline{C^j(t)}$, transition $(3)$ is executed and not $(2)$.
% %
% % The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
% % and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
% % and experienced a collision at time $t-1$
% Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).

% For player $j$ and round $t$, we now introduce a few events that are useful in the proof. First, for every $x=1,2,3,4,5$, we denote $I_x^j(t)$ the event that a transition of type $(x)$ occurs for player $j$ after the first $t$ observations (\ie, between round $t$ and round $t+1$, to decide $A^j(t+1)$).
% Formally they are defined by
% \begin{align*}
%   I_1^j(t) &:= \left(\overline{s^j(t)},\overline{C_j(t)}, A^j(t) \in \TopM(t)\right), \\
%   % go sitted
%   I_2^j(t) &:= \left(\overline{s^j(t)},C_j(t), A^j(t) \in \TopM(t)\right),
%   \ \ \ &\text{and} \ \ \
%   I_3(t) &:= \left(\overline{s^j(t)},A^j(t) \notin \TopM(t)\right), \\
%   % if sitted
%   I_3(t) &:= \left(s^j(t),A^j(t) \in \TopM(t)\right),
%   \ \ \ &\text{and} \ \ \
%   I_5(t) &:= \left(s^j(t),A^j(t) \notin \TopM(t)\right).
% \end{align*}
% %
% Then, we introduce $\widetilde{C^j}(t)$ as the event that a collision occurs for player $j$ at round $t$ if she is not yet fixed on her arm, that is
% \begin{equation}
%     \widetilde{C^j}(t) := \left(C^j(t), \overline{s^j(t)}\right).
% \end{equation}
